---
title: "cyclotomic"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

*The field of cyclotomic numbers.* 

___

The set of cyclotomic numbers is a field obtained by extending the set of 
rational numbers with the complex roots of unity. The main function used to 
construct a cyclotomic number in this package is `e`, it returns a principal 
root of the unity. For example `e(4)` is the principal fourth root of unity, 
that is the imaginary unit.

```{r}
library(cyclotomic2)
im <- e(4)
im^2
```

Arithmetic on cyclotomic numbers can be performed with this package. 
It is exact. In particular it allows to deal with the Gaussian rational numbers: 
the complex numbers whose both real and imaginary part are rational.

```{r}
a <- as.cyclotomic(5)
b <- as.cyclotomic("3/2")
(a + im * b)^2
```

Note that while `e(4)` is printed as `e(4)`, this is not the case for all 
roots of unity:

```{r}
e(9)
```

The set of cyclotomic numbers contains all the square roots of rational numbers, 
and therefore the package allows exact calculations on such square roots. For 
example, using float numbers, the following equality does not hold true:

```{r}
sqrt(5/3) == sqrt(5) / sqrt(3)
```

But it holds true using the cyclotomic arithmetic:

```{r}
cycSqrt("5/3") == cycSqrt(5) / cycSqrt(3)
```

The set of cyclotomic numbers also contains the cosine and sine of the rational 
multiples of pi. In particular, it contains the cosine and sine of any rational 
number when this number represents an angle given in degrees.

```{r}
cosDeg(60)
sinDeg(60) == cycSqrt(3) / 2
```

